\documentclass[12pt,letterpaper]{report}
%\usepackage{t1enc}
%\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\pagestyle{empty}
\textwidth6.15in
\oddsidemargin-.15in
\textheight9.5in
\topmargin-.75in

\def\ds{\displaystyle}
\def\ts{\textstyle}
\def\Z{{\bf Z}}
\def\Q{{\bf Q}}
\def\R{{\bf R}}
\def\P{{\cal P}}
\def\PS{\P(S)}
\def\iff{\Longleftrightarrow}
\def\implies{\Longrightarrow}
\def\nd{\mbox{ and }}
\def\o{\mbox{ or }}
\def\nt{\mbox{not }}
\def\gcd{\mbox{gcd}}

\begin{document}

\begin{center}
Foundations of Mathematics \\
Evaluating and Revising Proofs\\
\end{center}

Here are some attempted proofs, adapted by a colleague from past FOM student answers, of the following statement:
\begin{quote}
Suppose~$n$ is a positive integer.  If~$n^2$ is a multiple of~$3$, then~$n$ is a multiple of~$3$.
\end{quote} 

For each of the following proofs:
\begin{itemize}
\item  Read through the proof line by line.  Does each line follow from a known mathematical fact or from previous lines?
\item  Decide whether it is a valid proof.  Why do you think it is valid or invalid?
\item  Grade the proof on a scale from~$1$ (does not verify the statement at all) to~$5$ (a completely clear, convincing, and logically sound argument).
\item  What advice would you give the proof writer on how to change the proof?
\end{itemize}

\bigskip

\begin{enumerate}

\item  Let $n$ be an integer such that $n^2=3x$ where $x$ is an integer. Then $3 | n^2$. Since $n^2 = 3x$,  $nn = 3x$. Thus $3 | n$. Therefore if $n^2$ is a multiple of 3, then $n$ is a multiple of 3.

\item  Assume that $n^2$ is an odd positive integer that is divisible by 3. That is $n^2 = (3n+1)^2 = 9n^2 + 6n +1 = 3n(n+2)+1$. Therefore, $n^2$ is divisible by 3. Assume that $n^2$ is even and a multiple of 3. That is, $n^2 = (3n)^2 = 9n^2 = 3n(3n)$. Therefore, $n^2$ is a multiple of 3. If we factor $n^2 = 9n^2$, we get $3n(3n)$, which means that $n$ is a multiple of 3.

\item  Let $n$ be a positive integer such that $n^2$ is a multiple of 3. Then $n=3m$ where $m \in \Z$. So $n^2 = (3m)^2 = 9 m^2 = 3 (3m^2).$ This breaks down into $3m$ times $3m$ which shows that $m$ is a multiple of 3. 

\item  Suppose $n$ is not a multiple of 3. We will let $3k$ be a positive integer that is a multiple of 3, so that $3k+1$ and $3k+2$ are integers that are not multiples of 3. Now $n^2 = (3k+1)^2 = 9k^2 + 6k +1 = 3(3k^2+2k) + 1$. Since $ 3(3k^2+2k)$ is a multiple of 3, $ 3(3k^2+2k)+1$ is not. For the other case, $n^2 = (3k+2)^2 = 9k^2 + 12k +4 = 3(3k^2+4k+1) + 1$, which is not a multiple of 3. %Because $n^2$ is not a multiple of 3, we have a contradiction.

\bigskip
\bigskip
\bigskip

%\item Now that you've read and examined all of these proofs, work with your classmates to write a proof of your own, keeping the principles of good proof writing in mind.

\end{enumerate}

\end{document}